Compositional equivalences based on Open pNets

### 2021-10-07 2016-Weak-FH-Bisim/WeakBisim.tex
improved todos
changt otimes + typos mineurs

%\usepackage{amsfonts}
%\usepackage{fancyhdr}
%\usepackage{cite}
%\usepackage{ifthen}
%\usepackage{amssymb}
%\usepackage{fancyhdr}
%\usepackage{pifont}
\usepackage{stmaryrd}
\usepackage{mathtools,mathpartir}
\usepackage{proof}
%\usepackage{setspace}
%\usepackage{indentfirst}
\usepackage{amsmath,amssymb,amscd,mathrsfs}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
\usepackage{epsfig,color,subfigure,enumitem,soul}
\newcommand{\TODO}[1]{\textcolor{red}{\textbf{[TODO:#1]}}}
\newcommand{\NOTE}[1]{\textcolor{blue}{\textbf{[NOTE:#1]}}}
\newcommand{\ERIC}[1]{\textcolor{blue}{#1}}
\definecolor{darkgreen}{rgb}{0.1, 0.5, 0.1}
\newcommand{\LUDO}[1]{\textcolor{darkgreen}{#1}}
\newcommand{\RAB}[1]{\textcolor{magenta}{#1}}
\newcommand{\coloncolon}{{:\hspace{-.2ex}:}}
\newcommand{\bigcupdot}{\charfusion[\mathop]{\bigcup}{\cdot}}
\makeatletter
\newcommand{\raisemath}[1]{\mathpalette{\raisem@th{#1}}}
\newcommand{\raisem@th}[3]{\raisebox{#1}{$#2#3$}}
\makeatother
\newcommand{\shortotimes}{\!\otimes\!}
\usepackage{macrospNets}

@@

We develop here a semantics for a model of interacting processes with parameters and holes. 
Our approach is originaly inspired from Structured Operational Semantics with conditional premisses as in \cite{GROOTE1993263,VANGLABBEEK2004229}.
But we aim at a more constructive and implementable appraoch to compute the semantics (intuitively transitions including first order predicates) and to 
check equivalences for these open systems.
The main interest of our symbolic approach is to define a method to
prove properties directly on open structures; these properties will then be preserved
by any correct instantiation of the holes.
As a consequence, our model allows us to reason on composition
operators as well as on full-size distributed systems. The parametric
nature of the model and the properties of compositionability of the
equivalence relations are thus  the main
strengths of  our approach.

@@

This article also uses substitutions. Applying a substitution inside a term $t$ is denoted $t\subst{y_i\gets e_i}^{i\in I}$ and consists in replacing in parallel all the occurrences of variables $y_i$ in the term $t$ by the terms $e_i$. Note that a substitution is defined by a partial function that is applied on the variables inside a term. We let $\Post$ range over partial functions that are used as substitution and use the notation $\{y_i\gets e_i\}^{i\in I}$ to define such a partial function\footnote{When using this notation, we suppose, without loss of generality that each $y_i$ is different.}. These partial functions are sometimes called \emph{substitution functions} in the following. Thus,
 $\subst{\Post}$ is  the operation that applies,  in a parallel manner,  the substitution defined by the partial function $\Post$. $\otimes$ is a composition operator on these partial functions, such that for any term $t$ we have: $t\subst{\Post\shortotimes\Post'} = (t\subst{\Post'})\subst{\Post}$.
This property must also be valid when the substitution does not operate on all variables.
We thus define a composition operation as follows: 
%\TODO{version d eric reprise et corrigee}
%$(x_k\gets e_k)^{k\in K}\shortotimes \Post =  (x_k\gets e_k\subst{\Post})^{k\in K}$
%\ERIC{J'ai un probleme ici... quand $Post'$ agit sur des vars de $t$ qui ne sont pas substitutees par $Post$, alors les 2 defs de $\shortotimes$ ci-dessus ne sont pas equivalentes:}\\
%$F(x,z)(\subst{x\gets z}\shortotimes\subst{z\gets 1})=F(x,z)(\subst{x\gets 1})=F(1,z)$\\
%\ERIC{Alors que}\\
%$(F(x,z)\subst{x\gets z})\subst{z\gets 1})=F(z,z)\subst{z\gets 1}=F(1,1)$
%
%\ERIC{Une definition plus correcte serait peut-etre:}\\

\[(x_{k}\gets e_{k})^{k\in K}\shortotimes (x'_{k'}\gets e'_{k'})^{k'\in K'} =  
(x_{k}\gets e_{k}\subst{(x'_{k'}\gets e'_{k'})^{k'\in K'}})^{k\in K} \cup (x'_{k'}\gets e'_{k'})^{k'\in K''}\]
where $K''=\{k'\in K'|x'_{k'}\not\in\{x_k\}^{k\in K}\}$
%
%\LUDO{En fait comme on supose qu il y a pas de variable globale et donc que les post doivent affecter toutes les variables ce contre-exemple n est pas applicable! apres on peut definir comme ca si on veut ceinture et bretelles mais ca alourdit un peu la definitions}
%
%\ERIC{Heu Non... 
%  En pratique, dans la composition de posts de la regle WT3, les posts assignent bien toutes les variables de leur OT (strong), mais dans la composition, on fini par avoir des trous... Voir page 21-22. On est tombes tous les deux separemment, Rabea et moi, sur ce genre de cas.}
% 
%\LUDO{euuh ca doit etre lie qu fait qu il manque des ec leftarrowec dans les selfloop de l'OA de spec. Plutot que changer la def et les preuves je propose de rajouter la condition $vars(s')=dom(Post)$ pour tous les OAs ce qui ne change rien a part pour les selfloop que l on s autorisait apparemment a ecrire en omettant $x:=x$ ... au pire l'identite est un post toujours correct et valide}

@@

\begin{definition}[pNets]\label{def-pnets}
A pNet $\pNet$ is a hierarchical structure where leaves are pLTSs and holes:\\
$\pNet\triangleq \pLTS~|~\mylangle \pNet_i^{i\in I}, \Sort_j^{j\in J}, \symb{SV}_k^{k\in 
K}\myrangle$
where
\begin{itemize}
\item[$\bullet$] $I$ is a set of indices and $\pNet_i^{i\in I}$ is the family of sub-pNets indexed over $I$. $vars(P_i)$ and $vars(P_j)$ must be disjoint for $i\neq j$.
%  $\pNet_i^{i\in I}$ is a family of sub-pNets where $I\in\I_\P$ is the set over which sub-pNets are indexed.

@@

Figure \ref{SimpleProt:Spec} (right) shows a graphical view of the pNet $\symb{SimpleProtocolSpec}$ that specifies the system. The pNet is made of the composition of two pNets: a $\symb{SimpleSystem}$ node, and a $\symb{PerfectBuffer}$ sub-pNet. 
The full system implementation should be equivalent (e.g.  weakly bisimilar)
to thhis $\symb{SimpleProtocolSpec}$.  
The pNet has a tree-like structure.  The root node of the tree {\it SimpleSystem} is the top level of the pNet structure. It acts as the parallel operator. It consists of three nodes: two holes $P$ and $Q$ and one sub-pNet, denoted {\it PerfectBuffer}. Nodes of the tree are synchronized using four synchronization vectors, that  express the possible synchronizations between the parameterised actions of a subset of the nodes. For instance, in the vector $< \texttt{p-send(m)},\texttt{in(m)},\_> \rightarrow \texttt{\nounderline{in(m)}}$ only $P$ and {\it PerfectBuffer} nodes are involved in the synchronization. The synchronization between these processes occurs when  process $P$  performs $\texttt{p-send(m)}$  action sending a message, and  the {\it PerfectBuffer} accepts the message through an $\texttt{in(m)}$ action  at the same time; the result that will be returned at upper level is the action  $\texttt{\nounderline{in(m)}}$.

@@

For all predicate $\Pred$, for all partial functions $\Post$ such that $V\cap \dom(\Post)=\emptyset$. We have: 
		 \begin{mathpar}
    \openrule
         {
           \set{\beta}, \Pred\,',\Post\,'}
          {t \OTarrow {\alpha} t'}\in\mathcal{T}
\implies
    \openrule
         {
           \set{\beta}\subst{\Post}, \Pred\,'\subst{\Post}\land\Pred,\Post\shortotimes\Post\,'}
         {\raisemath{-2pt}{t \OTarrow {\alpha\subst{\Post}} {t'}}{}}
 \in\mathcal{T}
\end{mathpar}
\end{definition}

@@

The parametric nature of the automata entails a second original aspect of FH-bisimulation : the nature of the bisimulation relation itself.
 A classical relation between states can be seen as a function mapping pairs of state to a boolean value (true if the states are related, false if they are not). A FH-bisimulation relation maps pairs of states to boolean expressions that use variables of the two systems.  %Each variable used in $\Pred_{s,t}$ should refer to a state variable of $s$ or $t$ and there must be no ambiguity on the referenced variable, we thus require that the two compared automata have disjoint state variables.
Formally, a relation over the states of two open automata  $\mylangle J,\mathcal{S}_1, s_0,V_1,
   \mathcal{T}_1 \myrangle$ and $\mylangle J,\mathcal{S}_2,t_0,V_2, \mathcal{T}_2 \myrangle$ has the signature $\mathcal{S}_1\times\mathcal{S}_2\to\AlgB$. 
We suppose without loss of generality that the variables of the two open automata are disjoint.
We adopt a notation similar to standard relations and denote it
 $\mathcal{R}=\{(s,t|\Pred_{s,t})\}$, where: 1) For any pair $(s,t)\in \mathcal{S}_1\times \mathcal{S}_2$, there is a 
   single
      $(s,t|\Pred_{s,t})\in\mathcal{R}$  stating that $s$ and $t$ are related 
      if $\Pred_{s,t}$       is 
      True, i.e. the states are related when the value of the automata variables% in $s$ and $t$
  verify the predicate $\Pred_{s,t}$. 2) The free variables of  $\Pred_{s,t}$ belong to $V_1$ and $V_2$, i.e. $\vars(\Pred_{s,t})\subseteq V_1\cup V_2$.
 FH-bisimulation is defined formally: 
 \begin{definition}[Strong FH-bisimulation]\label{def-FH-bisim} ~\\
\noindent
\begin{minipage}{0.67\linewidth} 	Suppose 
   $A_1 = \mylangle J,\mathcal{S}_1, s_0,V_1,
   \mathcal{T}_1 \myrangle$ and $A_2 = \mylangle J,\mathcal{S}_2,t_0,V_2, \mathcal{T}_2 \myrangle$
   are open automata with identical holes of the same sort, with disjoint sets of variables ($V_1\cap V_2=\emptyset$).

@@

We are now able to build a weak open automaton from an open automaton. This is done in a way that resembles the process of $\tau$ saturation: we add  $\tau$ open transitions before or after another (observable or not) open transition.
\begin{definition}[Building a weak open automaton]\label{def:buildweakOT}
  Let $A = \mylangle J,\mathcal{S},s_0,V,\mathcal{T}\myrangle$ be an open automaton. 
The weak open automaton \emph{derived} from $A$ is an open automaton  $\mylangle J,\mathcal{S},s_0,V,\WT\myrangle$ where $\WT$ is derived from $\mathcal{T}$ by saturation, applying the following rules:
%\noindent{\bf Invisible $\tau$ transitions:}
\begin{mathpar}
 \openrule
         {
           \emptyset,\symb{True},\Id(V)}
         {s \OTWeakarrow {\tau} s} \in \WT  \qquad \WTUn
 \end{mathpar}
and
\begin{mathpar}
\inferrule{
 \openrule
         {
           \set{\beta},\Pred,\Post}
         {s \OTarrow {\alpha} s'} \in \mathcal{T}
} 
{ \openrule
         {
           \vis{\set{\beta}}\!,\Pred,\Post
				 } {s \OTWeakarrow {\alpha} s'} \in \WT
}\qquad \WTDeux
%\inferrule{
% \openrule
%         {
%           \set{\beta},\Pred,\Post}
%         {s \OTarrow {\tau} s'} \in \mathcal{T}
%\\
% \openrule
%         {
%           \set{\gamma},\Pred\,',({x_k\gets e_k})^{k\in K}   }
%         {s' \OTWeakarrow {\tau} s''} \in \WT
%}
%{ \openrule
%         {
%           \vis{\set{\beta}}\dotcup\set{\gamma},\Pred\land\Pred\,'\subst{\Post\,},
%				({x_k\gets (e_k\subst{\Post\,})})^{k\in K} } 
%         {s \OTWeakarrow {\tau} s''} \in \WT
%}
 \end{mathpar}
 and
%\noindent{\bf Other transitions:}
\begin{mathpar}
\inferrule {\openrule
         {
           \set{\gamma_1},\Pred_1,\Post_1   }
         {s \OTWeakarrow {\tau} s_1} \in \WT
\qquad
\openrule
         {
           \set{\gamma_2},\Pred_2,\Post_2  }
         {s_1 \OTWeakarrow {\alpha} s_2} \in \WT
\qquad
\openrule
         {
           \set{\gamma_3},\Pred_3,\Post_3  }
         {s_2 \OTWeakarrow {\tau} s'} \in\WT
\\
\Pred=\Pred_1\land\Pred_2\subst{\Post_1}\land \Pred_3\subst{\Post_2\shortotimes\Post_1}
\\
\set{\gamma}=\set{\gamma_1}\dotcup \set{\gamma_2}\subst{\Post_1}\dotcup\set{\gamma_3}\subst{\Post_2\shortotimes \Post_1}\\
\alpha'=\alpha\subst{\Post_1}
}
{
\openrule
         {\set{\gamma}
           ,
		\Pred,
				\Post_3\shortotimes\Post_2\shortotimes\Post_1} 
         {s \OTWeakarrow {\alpha'} s'} \in\WT
} \WTTrois
\end{mathpar}

\end{definition}
Rule $\WTUn$ states that it is always possible to do a non-observable transition, where the state is unchanged and the holes perform no action. Rule~$\WTDeux$ states that each open transition can be considered as a weak open transition. The last rule is the most interesting:  Rule~$\WTTrois$ allows any number of $\tau$ transitions before or after any weak open transition. This rules carefully composes predicates, effects, and actions of the holes, indeed in the rule, predicate $\Pred_2$ manipulates variables of $s_1$ that result from the first weak open transition. Their values thus depend on the initial state but also on the effect (as a substitution function $\Post_1$) of the first weak open transition. In the same manner, $\Pred_3$ must be applied the substitution defined by the composition $\Post_2\shortotimes\Post_1$. Similarly, effects on variables must be applied to obtain the global effect of the composed weak open transition, it must also be applied to observable actions of the holes, and to the global action of the weak open transition.

@@

\begin{example} [A weak open-transition]
Figure \ref{WOT2} shows the construction of one of the weak transitions of the open automaton of \symb{SimpleProtocolSpec}. On the top we show the subset of the original open automaton (from Figure \ref{SimpleProtCounter:SpecOA}) considered here, and at the bottom the generated weak transition.  For readability, we abbreviate the weak open transitions encoded by $\openrule   {\{\}, True,	() } {s \OTWeakarrow \tau s'}$  as $W_\tau$. The weak open transition shown here is the transition delivering the result of the algorithm to hole $Q$ by applying rules: \WTUn,\WTDeux, and \WTTrois. First rule \WTUn~ adds a $WT_\tau$ loop on each state. Rule \WTDeux~ transforms each 3 OTs into WOTs.   Then consider application of Rule \WTTrois~ on a sequence 3  WOTs.   $\openrule
         {\{\}, True,
			(\texttt{b\_ec}\gets \texttt{b\_ec}+1) }
         {\texttt{b1}\OTWeakarrow \tau \texttt{b1}}$; $\openrule
         {\{\}, True,
			(\texttt{b\_ec}\gets \texttt{b\_ec}+1) }
         {\texttt{b1} \OTWeakarrow \tau \texttt{b1}}$;  $\openrule
         {\{\}, True,
			() }
         {\texttt{b1} \OTWeakarrow \tau \texttt{b1}}$. Then Rule \WTTrois~ produces  $\openrule
         {\{\}, True,
			(\texttt{b\_ec}\gets \texttt{b\_ec}+2) }
         {\texttt{b1} \OTWeakarrow \tau \texttt{b1}}$. We can iterate this construction an arbitrary number of times, getting for any natural number $n$ a weak open transition:
  $\openrule
         {\emptyset, True,
			(\texttt{ec}\gets \texttt{ec}+n) }
        {\texttt{b1} \OTWeakarrow \tau \texttt{b1}}\, \forall n \geq 0$.  Finally,  applying again \WTTrois, and using the central open transition having \texttt{\nounderline{out(b\_msg,b\_ec)}}  as $\alpha$, we get the resulting weak open transition between b1 and b0 (as shown in Figure \ref{WOT2}).  Applying the substitutions finally yields the weak transitions family $WS_7$ in Figure  \ref{SimpleProtCounter:WeakSpecOA}.

@@

Then the effect of one single $100 \xrightarrow{OT_4} 210 \xrightarrow{OT_5} 220 \xrightarrow{OT_6} 100$ loop is\footnote{when showing  the result of $Posts$ composition, we will omit the identity substitution functions introduced by the $\shortotimes$ definition in page \pageref{def:substitutions}}:
\[post_{456} = post_6 \shortotimes\ post_5 \shortotimes\ post_4
= (\texttt{s\_ec}\gets \texttt{s\_ec}+1)\]

So if we denote ${post_{456*}}$ any iteration of this loop, we get ${post_{456*}} = (\texttt{s\_ec}\gets \texttt{s\_ec}+n)$ for any $n\ge 0$, and the \Post~ of the weak OT  $WI_{3}$ is:\\
 $Post_3 = post_{456*}\shortotimes\ post_3 = (\texttt{s\_msg}\gets m, \texttt{s\_ec}\gets n), \forall n\ge 0$ and \Post~ of  $WI_{3a}$  is:\\ $post_4\shortotimes {post_{456*}}\shortotimes\ post_3 = (\texttt{m\_msg}\gets m, \texttt{m\_ec}\gets n), \forall n\ge 0$.
\medskip

@@

Two pNets are weak FH-bisimilar if there exists a relation between their associated 
automata that is a Weak FH-bisimulation and their initial states are in the relation, i.e. 
the predicate associated to the relation between the initial states is \True.
 \end{definition}

@@

Our first important result is that Weak FH-bisimilarity is an equivalence in the same way as strong FH-bisimilarity:

>>>>

%\usepackage{amsfonts}
%\usepackage{fancyhdr}
%\usepackage{cite}
%\usepackage{ifthen}
%\usepackage{amssymb}
%\usepackage{fancyhdr}
%\usepackage{pifont}
\usepackage{stmaryrd}
\usepackage{mathtools,mathpartir}
\usepackage{proof}
%\usepackage{setspace}
%\usepackage{indentfirst}
\usepackage{amsmath,amssymb,amscd,mathrsfs}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
\usepackage{epsfig,color,subfigure,enumitem,soul}
\newcommand{\TODO}[1]{\textcolor{red}{\textbf{[TODO:#1]}}}
\newcommand{\NOTE}[1]{\textcolor{blue}{\textbf{[NOTE:#1]}}}
\newcommand{\ERIC}[1]{\textcolor{blue}{#1}}
\definecolor{darkgreen}{rgb}{0.1, 0.5, 0.1}
\newcommand{\LUDO}[1]{\textcolor{darkgreen}{#1}}
\newcommand{\RAB}[1]{\textcolor{magenta}{#1}}
\newcommand{\coloncolon}{{:\hspace{-.2ex}:}}
\newcommand{\bigcupdot}{\charfusion[\mathop]{\bigcup}{\cdot}}
\makeatletter
\newcommand{\raisemath}[1]{\mathpalette{\raisem@th{#1}}}
\newcommand{\raisem@th}[3]{\raisebox{#1}{$#2#3$}}
\makeatother
\newcommand{\shortotimes}{\!\otimes\!}
\newcommand{\shortodot}{\!\odot\!}
\usepackage{macrospNets}

@@

We develop here a semantics for a model of interacting processes with parameters and holes. 
Our approach is originally inspired from Structured Operational Semantics with conditional premisses as in \cite{GROOTE1993263,VANGLABBEEK2004229}.
But we aim at a more constructive and implementable approach to compute the semantics (intuitively transitions including first order predicates) and to 
check equivalences for these open systems.
The main interest of our symbolic approach is to define a method to
prove properties directly on open structures; these properties will then be preserved
by any correct instantiation of the holes.
As a consequence, our model allows us to reason on composition
operators as well as on full-size distributed systems. The parametric
nature of the model and the properties of compositionability of the
equivalence relations are thus  the main
strengths of  our approach.

@@

This article also uses substitutions. Applying a substitution inside a term $t$ is denoted $t\subst{y_i\gets e_i}^{i\in I}$ and consists in replacing in parallel all the occurrences of variables $y_i$ in the term $t$ by the terms $e_i$. Note that a substitution is defined by a partial function that is applied on the variables inside a term. We let $\Post$ range over partial functions that are used as substitution and use the notation $\{y_i\gets e_i\}^{i\in I}$ to define such a partial function\footnote{When using this notation, we suppose, without loss of generality that each $y_i$ is different.}. These partial functions are sometimes called \emph{substitution functions} in the following. Thus,
 $\subst{\Post}$ is  the operation that applies,  in a parallel manner,  the substitution defined by the partial function $\Post$. $\odot$ is a composition operator on these partial functions, such that for any term $t$ we have: $t\subst{\Post\shortodot\Post'} = (t\subst{\Post'})\subst{\Post}$.
This property must also be valid when the substitution does not operate on all variables.
We thus define a composition operation as follows: 
%\TODO{version d eric reprise et corrigee}
%$(x_k\gets e_k)^{k\in K}\shortotimes \Post =  (x_k\gets e_k\subst{\Post})^{k\in K}$
%\ERIC{J'ai un probleme ici... quand $Post'$ agit sur des vars de $t$ qui ne sont pas substitutees par $Post$, alors les 2 defs de $\shortotimes$ ci-dessus ne sont pas equivalentes:}\\
%$F(x,z)(\subst{x\gets z}\shortotimes\subst{z\gets 1})=F(x,z)(\subst{x\gets 1})=F(1,z)$\\
%\ERIC{Alors que}\\
%$(F(x,z)\subst{x\gets z})\subst{z\gets 1})=F(z,z)\subst{z\gets 1}=F(1,1)$
%
%\ERIC{Une definition plus correcte serait peut-etre:}\\

\[(x_{k}\gets e_{k})^{k\in K}\shortodot (x'_{k'}\gets e'_{k'})^{k'\in K'} =  
(x_{k}\gets e_{k}\subst{(x'_{k'}\gets e'_{k'})^{k'\in K'}})^{k\in K} \cup (x'_{k'}\gets e'_{k'})^{k'\in K''}\]
where $K''=\{k'\in K'|x'_{k'}\not\in\{x_k\}^{k\in K}\}$
%
%\LUDO{En fait comme on supose qu il y a pas de variable globale et donc que les post doivent affecter toutes les variables ce contre-exemple n est pas applicable! apres on peut definir comme ca si on veut ceinture et bretelles mais ca alourdit un peu la definitions}
%
%\ERIC{Heu Non... 
%  En pratique, dans la composition de posts de la regle WT3, les posts assignent bien toutes les variables de leur OT (strong), mais dans la composition, on fini par avoir des trous... Voir page 21-22. On est tombes tous les deux separemment, Rabea et moi, sur ce genre de cas.}
% 
%\LUDO{euuh ca doit etre lie qu fait qu il manque des ec leftarrowec dans les selfloop de l'OA de spec. Plutot que changer la def et les preuves je propose de rajouter la condition $vars(s')=dom(Post)$ pour tous les OAs ce qui ne change rien a part pour les selfloop que l on s autorisait apparemment a ecrire en omettant $x:=x$ ... au pire l'identite est un post toujours correct et valide}

@@

\begin{definition}[pNets]\label{def-pnets}
A pNet $\pNet$ is a hierarchical structure where leaves are pLTSs and holes\\
$\pNet\triangleq \pLTS~|~\mylangle \pNet_i^{i\in I}, \Sort_j^{j\in J}, \symb{SV}_k^{k\in 
K}\myrangle$
where:
\begin{itemize}
\item[$\bullet$] $I$ is a set of indices and $\pNet_i^{i\in I}$ is the family of sub-pNets indexed over $I$. $vars(P_i)$ and $vars(P_j)$ must be disjoint for $i\neq j$.
%  $\pNet_i^{i\in I}$ is a family of sub-pNets where $I\in\I_\P$ is the set over which sub-pNets are indexed.

@@

Figure \ref{SimpleProt:Spec} (right) shows a graphical view of the pNet $\symb{SimpleProtocolSpec}$ that specifies the system. The pNet is made of the composition of two pNets: a $\symb{SimpleSystem}$ node, and a $\symb{PerfectBuffer}$ sub-pNet. 
The full system implementation should be equivalent (e.g.  weakly bisimilar)
to thhis $\symb{SimpleProtocolSpec}$.  
The pNet has a tree-like structure.  The root node of the tree {\it SimpleSystem} is the top level of the pNet structure. It acts as the parallel operator. It consists of three nodes: two holes $P$ and $Q$ and one sub-pNet, denoted {\it PerfectBuffer}. Nodes of the tree are synchronised using four synchronisation vectors, that  express the possible synchronisations between the parameterised actions of a subset of the nodes. For instance, in the vector $< \texttt{p-send(m)},\texttt{in(m)},\_> \rightarrow \texttt{\nounderline{in(m)}}$ only $P$ and {\it PerfectBuffer} nodes are involved in the synchronisation. The synchronisation between these processes occurs when  process $P$  performs $\texttt{p-send(m)}$  action sending a message, and  the {\it PerfectBuffer} accepts the message through an $\texttt{in(m)}$ action  at the same time; the result that will be returned at upper level is the action  $\texttt{\nounderline{in(m)}}$.

@@

For all predicate $\Pred$, for all partial functions $\Post$ such that $V\cap \dom(\Post)=\emptyset$. We have: 
		 \begin{mathpar}
    \openrule
         {
           \set{\beta}, \Pred\,',\Post\,'}
          {t \OTarrow {\alpha} t'}\in\mathcal{T}
\implies
    \openrule
         {
           \set{\beta}\subst{\Post}, \Pred\,'\subst{\Post}\land\Pred,\Post\shortodot\Post\,'}
         {\raisemath{-2pt}{t \OTarrow {\alpha\subst{\Post}} {t'}}{}}
 \in\mathcal{T}
\end{mathpar}
\end{definition}

@@

The parametric nature of the automata entails a second original aspect of FH-bisimulation: the nature of the bisimulation relation itself.
 A classical relation between states can be seen as a function mapping pairs of state to a boolean value (true if the states are related, false if they are not). A FH-bisimulation relation maps pairs of states to boolean expressions that use variables of the two systems.  %Each variable used in $\Pred_{s,t}$ should refer to a state variable of $s$ or $t$ and there must be no ambiguity on the referenced variable, we thus require that the two compared automata have disjoint state variables.
Formally, a relation over the states of two open automata  $\mylangle J,\mathcal{S}_1, s_0,V_1,
   \mathcal{T}_1 \myrangle$ and $\mylangle J,\mathcal{S}_2,t_0,V_2, \mathcal{T}_2 \myrangle$ has the signature $\mathcal{S}_1\times\mathcal{S}_2\to\AlgB$. 
We suppose without loss of generality that the variables of the two open automata are disjoint.
We adopt a notation similar to standard relations and denote it
 $\mathcal{R}=\{(s,t|\Pred_{s,t})\}$, where: 1) For any pair $(s,t)\in \mathcal{S}_1\times \mathcal{S}_2$, there is a 
   single
      $(s,t|\Pred_{s,t})\in\mathcal{R}$  stating that $s$ and $t$ are related 
      if $\Pred_{s,t}$       is 
      True, i.e. the states are related when the value of the automata variables% in $s$ and $t$
  verify the predicate $\Pred_{s,t}$. 2) The free variables of  $\Pred_{s,t}$ belong to $V_1$ and $V_2$, i.e. $\vars(\Pred_{s,t})\subseteq V_1\cup V_2$.
 FH-bisimulation is defined formally: 
 \begin{definition}[Strong FH-bisimulation]\label{def-FH-bisim} ~\\
\noindent
\begin{minipage}{0.67\linewidth} 	Suppose 
   $A_1 = \mylangle J,\mathcal{S}_1, s_0,V_1,
   \mathcal{T}_1 \myrangle$ and $A_2 = \mylangle J,\mathcal{S}_2,t_0,V_2, \mathcal{T}_2 \myrangle$
   are open automata with identical holes of the same sort, with disjoint sets of variables ($V_1\cap V_2=\emptyset$).

@@

We are now able to build a weak open automaton from an open automaton. This is done in a way that resembles the process of $\tau$ saturation: we add  $\tau$ open transitions before or after another (observable or not) open transition.
\begin{definition}[Building a weak open automaton]\label{def:buildweakOT}
  Let $A = \mylangle J,\mathcal{S},s_0,V,\mathcal{T}\myrangle$ be an open automaton. 
The weak open automaton \emph{derived} from $A$ is an open automaton  $\mylangle J,\mathcal{S},s_0,V,\WT\myrangle$ where $\WT$ is derived from $\mathcal{T}$ by saturation, applying the following rules:
%\noindent{\bf Invisible $\tau$ transitions:}
\begin{mathpar}
 \openrule
         {
           \emptyset,\symb{True},\Id(V)}
         {s \OTWeakarrow {\tau} s} \in \WT  \qquad \WTUn
 \end{mathpar}
and
\begin{mathpar}
\inferrule{
 \openrule
         {
           \set{\beta},\Pred,\Post}
         {s \OTarrow {\alpha} s'} \in \mathcal{T}
} 
{ \openrule
         {
           \vis{\set{\beta}}\!,\Pred,\Post
				 } {s \OTWeakarrow {\alpha} s'} \in \WT
}\qquad \WTDeux
%\inferrule{
% \openrule
%         {
%           \set{\beta},\Pred,\Post}
%         {s \OTarrow {\tau} s'} \in \mathcal{T}
%\\
% \openrule
%         {
%           \set{\gamma},\Pred\,',({x_k\gets e_k})^{k\in K}   }
%         {s' \OTWeakarrow {\tau} s''} \in \WT
%}
%{ \openrule
%         {
%           \vis{\set{\beta}}\dotcup\set{\gamma},\Pred\land\Pred\,'\subst{\Post\,},
%				({x_k\gets (e_k\subst{\Post\,})})^{k\in K} } 
%         {s \OTWeakarrow {\tau} s''} \in \WT
%}
 \end{mathpar}
 and
%\noindent{\bf Other transitions:}
\begin{mathpar}
\inferrule {\openrule
         {
           \set{\gamma_1},\Pred_1,\Post_1   }
         {s \OTWeakarrow {\tau} s_1} \in \WT
\qquad
\openrule
         {
           \set{\gamma_2},\Pred_2,\Post_2  }
         {s_1 \OTWeakarrow {\alpha} s_2} \in \WT
\qquad
\openrule
         {
           \set{\gamma_3},\Pred_3,\Post_3  }
         {s_2 \OTWeakarrow {\tau} s'} \in\WT
\\
\Pred=\Pred_1\land\Pred_2\subst{\Post_1}\land \Pred_3\subst{\Post_2\shortodot\Post_1}
\\
\set{\gamma}=\set{\gamma_1}\dotcup \set{\gamma_2}\subst{\Post_1}\dotcup\set{\gamma_3}\subst{\Post_2\shortodot \Post_1}\\
\alpha'=\alpha\subst{\Post_1}
}
{
\openrule
         {\set{\gamma}
           ,
		\Pred,
				\Post_3\shortodot\Post_2\shortodot\Post_1} 
         {s \OTWeakarrow {\alpha'} s'} \in\WT
} \WTTrois
\end{mathpar}

\end{definition}
Rule $\WTUn$ states that it is always possible to do a non-observable transition, where the state is unchanged and the holes perform no action. Rule~$\WTDeux$ states that each open transition can be considered as a weak open transition. The last rule is the most interesting:  Rule~$\WTTrois$ allows any number of $\tau$ transitions before or after any weak open transition. This rules carefully composes predicates, effects, and actions of the holes, indeed in the rule, predicate $\Pred_2$ manipulates variables of $s_1$ that result from the first weak open transition. Their values thus depend on the initial state but also on the effect (as a substitution function $\Post_1$) of the first weak open transition. In the same manner, $\Pred_3$ must be applied the substitution defined by the composition $\Post_2\shortodot\Post_1$. Similarly, effects on variables must be applied to obtain the global effect of the composed weak open transition, it must also be applied to observable actions of the holes, and to the global action of the weak open transition.

@@

\begin{example} [A weak open-transition]
Figure \ref{WOT2} shows the construction of one of the weak transitions of the open automaton of \symb{SimpleProtocolSpec}. On the top we show the subset of the original open automaton (from Figure \ref{SimpleProtCounter:SpecOA}) considered here, and at the bottom the generated weak transition.  For readability, we abbreviate the weak open transitions encoded by $\openrule   {\{\}, True,	() } {s \OTWeakarrow \tau s'}$  as $W_\tau$. The weak open transition shown here is the transition delivering the result of the algorithm to hole $Q$ by applying rules: \WTUn,\WTDeux, and \WTTrois. First rule \WTUn~ adds a $WT_\tau$ loop on each state. Rule \WTDeux~ transforms each 3 OTs into WOTs.   Then consider application of Rule \WTTrois~ on a sequence 3  WOTs.   $\openrule
         {\{\}, True,
			(\texttt{b\_ec}\gets \texttt{b\_ec}+1) }
         {\texttt{b1}\OTWeakarrow \tau \texttt{b1}}$; $\openrule
         {\{\}, True,
			(\texttt{b\_ec}\gets \texttt{b\_ec}+1) }
         {\texttt{b1} \OTWeakarrow \tau \texttt{b1}}$;  $\openrule
         {\{\}, True,
			() }
         {\texttt{b1} \OTWeakarrow \tau \texttt{b1}}$. The result will be:   $\openrule
         {\{\}, True,
			(\texttt{b\_ec}\gets \texttt{b\_ec}+2) }
         {\texttt{b1} \OTWeakarrow \tau \texttt{b1}}$. We can iterate this construction an arbitrary number of times, getting for any natural number $n$ a weak open transition:
  $\openrule
         {\emptyset, True,
			(\texttt{ec}\gets \texttt{ec}+n) }
        {\texttt{b1} \OTWeakarrow \tau \texttt{b1}}\, \forall n \geq 0$.  Finally,  applying again \WTTrois, and using the central open transition having \texttt{\nounderline{out(b\_msg,b\_ec)}}  as $\alpha$, we get the resulting weak open transition between b1 and b0 (as shown in Figure \ref{WOT2}).  Applying the substitutions finally yields the weak transitions family $WS_7$ in Figure  \ref{SimpleProtCounter:WeakSpecOA}.

@@

Then the effect of one single $100 \xrightarrow{OT_4} 210 \xrightarrow{OT_5} 220 \xrightarrow{OT_6} 100$ loop is\footnote{when showing  the result of $Posts$ composition, we will omit the identity substitution functions introduced by the $\shortodot$ definition in page \pageref{def:substitutions}}:
\[post_{456} = post_6 \shortodot\ post_5 \shortodot\ post_4
= (\texttt{s\_ec}\gets \texttt{s\_ec}+1)\]

So if we denote ${post_{456*}}$ any iteration of this loop, we get ${post_{456*}} = (\texttt{s\_ec}\gets \texttt{s\_ec}+n)$ for any $n\ge 0$, and the \Post~ of the weak OT  $WI_{3}$ is:\\
 $Post_3 = post_{456*}\shortodot\ post_3 = (\texttt{s\_msg}\gets m, \texttt{s\_ec}\gets n), \forall n\ge 0$ and \Post~ of  $WI_{3a}$  is:\\ $post_4\shortodot {post_{456*}}\shortodot\ post_3 = (\texttt{m\_msg}\gets m, \texttt{m\_ec}\gets n), \forall n\ge 0$.
\medskip

@@

Two pNets are weak FH-bisimilar if there exists a relation between their associated 
automata that is a weak FH-bisimulation and their initial states are in the relation, i.e. 
the predicate associated to the relation between the initial states is \True.
 \end{definition}

@@

Our first important result is that weak FH-bisimilarity is an equivalence in the same way as strong FH-bisimilarity:
